### 2025-04-14 reports/paper/paper.qmd
prepare for revisions

Non-Hodgkin Lymphoma (NHL) incidence in the United States doubled
between 1973 and 1994 before plateauing at around 19 per 100,000 persons
per year, making it the seventh most common cancer in the country
[@seer_1994; @Ekstrom-Smedby_2006]. The strongest known risk factor of
NHL is immunosuppression, both congenital and acquired
[@Filipovich_1992; @Chiu_2015]. However, since immunodeficiency and
infection with HIV are rare, they cannot fully explain the historic rise
or present burden of NHL [@Shiels_2013]. The rise in NHL incidence
coincided with a period of rapid and extensive chemicalization in
industry, agriculture and warfare; environmental and occupational
exposures may play an important explanatory role in the epidemiology of
NHL [@Nelson_2005; @Romero_2021].

@@

Common approaches for addressing violations in positivity are
model-based smoothing when using outcome regression methods, weight
truncation when using propensity methods, or a combination of both.
Another approach is to only consider causal contrasts that are supported
by the observed data, *i.e.,* those involving hypothetical interventions
with propensity score strictly greater than zero. Here, we defined
hypothetical supportable interventions on exposure to soluble MWF in the
United Auto Workers-General Motors (UAW-GM) occupational cohort and
estimated their effects on NHL incidence between 1985 and 2015 relative
to no exposure intervention while guaranteeing positivity.

@@

We estimated the cumulative incidence of NHL between 1985 and 2015 under
supportable interventions based on selected target exposure limits on
annual average daily exposure to soluble MWF by applying the
hazard-extended iterative conditional expectation (ICE) parametric
g-formula [@Wen_2020]. First, we estimated the expected number of NHL
cases that we would observe if there were no censoring and no limit on
exposure. Then, we contrasted this case count to that under supportable
interventions based on five hypothetical target exposure limits and no
censoring. The five target exposure limits were (1) 2.0, (2) 1.0, (3)
0.5, (4) 0.25, and (5) 0.05 mg/m^3^. The National Institute for
Occupational Safety and Health (NIOSH) recommended exposure limit (REL)
for time-weighted average total particulate mass (PM) composed of MWF is
0.5 mg/m^3^ [@niosh_1998].

>>>>

Non-Hodgkin Lymphoma (NHL) incidence in the United States doubled
between 1973 and 1994 before plateauing at around 19 per 100,000
persons per year, making it the seventh most common cancer in the
country [@seer_1994; @Ekstrom-Smedby_2006]. The strongest known risk
factors of NHL are autoimmune disease and immunosuppression
[@Filipovich_1992; @Chiu_2015]. The relationship between autoimmune
diseases NHL is specific in that each disease is associated with a
particular subtype of NHL [@Thandra_2021]. Immunosuppression and
infection with HIV are strongly associated with Burkitt lymphoma
[@Thandra_2021]. These risk factors cannot fully explain the historic
rise or present burden of NHL, however [@Shiels_2013]. Since the
rise in NHL incidence coincided with a period of rapid and extensive
chemicalization in industry, agriculture and warfare; environmental
and occupational exposures may play an important explanatory role in
the epidemiology of NHL [@Nelson_2005; @Romero_2021]. Existing
epidemiologic studies suggest that occupational exposures may increase
the risk of NHL without subtype specificity [@Thandra_2021].

@@

When positivity is not satisfied, causal inference methods that rely
only on models for the counterfactual outcome, and not the treatment
mechanism, may still yield valid causal estimates, if the models are
correctly specified. We evaluated how hypothetical interventions on
exposure to soluble MWF would affect the incidence of NHL between 1985
and 2015 in the United Auto Workers-General Motors (UAW-GM)
occupational cohort using the iterative conditional expectation (ICE)
parametric g-formula, which relies on models for the outcome only. To
further address potential violations in positivity, we defined and
evaluated hypothetical interventions resulting in post-intervention
exposure histories whose propensity score is strictly greater than 
zero _i.e._, exposure histories that can be found in the observed
data. We call these hypothetical interventions "supportable".

@@

We estimated the cumulative incidence of NHL between 1985 and 2015 under
supportable interventions based on selected target exposure limits on
annual average daily exposure to soluble MWF by applying the
hazard-extended ICE parametric  we estimated the expected number of NHL
cases that we would observe if there were no censoring and no limit on
exposure. Then, we contrasted this case count to that under supportable
interventions based on five hypothetical target exposure limits and no
censoring. The five target exposure limits were (1) 2.0, (2) 1.0, (3)
0.5, (4) 0.25, and (5) 0.05 mg/m^3^. The National Institute for
Occupational Safety and Health (NIOSH) recommended exposure limit (REL)
for time-weighted average total particulate mass (PM) composed of MWF is
0.5 mg/m^3^ [@niosh_1998].
